\input common
\def\begthm#1#2{
\vbox\bgroup
{\headerfont {Theorem #1:} \quad #2}
\smallskip\hrule\smallskip
}

\def\endthm{
\egroup
\bigskip\bigskip
}

\begthm{3.1}{Differentiable Implies Continuous}
If $f$ is differentiable at $a$, then $f$ is continuous at $a$.
\endthm

\begthm{3.1a}{Not Continuous implies Not Differentiable}
If $f$ is not continuous at $a$, then $f$ is not differentiable
at $a$.
\endthm

\begthm{3.2}{Constant Rule}
If $c$ is a real number, then ${d \over dx}(c) = 0$.
\endthm

\begthm{3.3}{Power Rule}
If $n$ is a nonnegative integer, then ${d \over dx}(x^n) = nx^{n - 1}$
\endthm

\begthm{3.4}{Constant Multiple Rule}
If $f$ is differentiable at $x$ and $c$ is a constant, then
$$\
{d \over dx}(cf(x)) = c f'(x).
$$
\endthm

\begthm{3.5}{Sum Rule}
If $f$ and $g$ are differentiable at $x$, then

$$\
{d \over dx}(f(x) + g(x)) = f'(x) + g'(x).
$$
\endthm

\begthm{3.6}{The Derivative of $e^x$}
The function $f(x) = e^x$ is differentiable for all real numbers $x$,
and
$$\
{d \over dx}(e^x) = e^x
$$
\endthm

\begthm{3.7}{Product Rule}
If $f$ and $g$ are differentiable at $x$, then

$$\
{d \over dx}(f(x)g(x)) = f'(x)g(x) + f(x)g'(x).
$$
\endthm

\begthm{3.8}{Quotient Rule}
If $f$ and $g$ are differntiable at $x$ and $g(x) \ne 0$,
then the derivative of $f / g$ at $x$ exists and
$$\
{d \over dx}\bigg({f(x) \over g(x)}\bigg) = {
g(x)f'(x) - f(x)g'(x)
\over
(g(x))^2
}
$$
\endthm

\begthm{3.9}{Power Rule (general form)}
If $n$ is any real number, then
$$\
{d \over dx}(x^n) = nx^{n - 1}
$$
\endthm

\begthm{2.1}{Relationship between One-Sided and Two-Sided Limits}
Assume $f$ is defined for all $x$ near $a$ except possibly
at $a$. Then

$$\
\lim_{x \to a} f(x) = L
$$

if and only if

$$\
\lim_{x \to a^{-}} = L
$$

and

$$\
\lim_{x \to a^{+}} = L
$$
\endthm

\begthm{2.2}{Limits of a Linear Function}
Let $a$, $b$, and $m$ be real numbers. For linear functions
$f(x) = mx + b$,

$$\
\lim_{x \to a} {f(x)} = f(a) = ma + b.
$$
\endthm

\begthm{2.3a}{Limit Law: Sum}
$$\
\lim_{x \to a}{
(f(x) + g(x))
} =
\lim_{x \to a}{f(x)} +
\lim_{x \to a}{g(x)} 
$$
\endthm

\begthm{2.3b}{Limit Law: Difference}
$$\
\lim_{x \to a}{
(f(x) - g(x))
} =
\lim_{x \to a}{f(x)} -
\lim_{x \to a}{g(x)} 
$$
\endthm

\begthm{2.3c}{Limit Law: Constant multiple}
$$\
\lim_{x \to a}{
cf(x)
} =
c\lim_{x \to a}{f(x)}
$$
\endthm

\begthm{2.3d}{Limit Law: Product}
$$\
\lim_{x \to a}{
(f(x)g(x))
} =
\bigg(
\lim_{x \to a}{f(x)}
\bigg)
\bigg(
\lim_{x \to a}{g(x)} 
\bigg)
$$
\endthm

\begthm{2.3e}{Limit Law: Quotient}
$$\
\lim_{x \to a}{
\bigg(
{
f(x)
\over
g(x)
}
\bigg)
} =
{
{\lim_{x \to a}{f(x)}}
\over
{\lim_{x \to a}{g(x)}}
}
$$

provided

$$\
\lim_{x \to a}{g(x)} \ne 0
$$
\endthm

\begthm{2.3f}{Limit Law: Power}
$$\
\lim_{x \to a}{
(f(x))^n
}
=
\big(
\lim_{x \to a}{
(f(x))
}
\big)^n
$$
\endthm

\begthm{2.3g}{Limit Law: Root}
$$\
\lim_{x \to a}{
(f(x))^{1/n}
}
=
\big(
\lim_{x \to a}{
(f(x))
}
\big)^{1/n}
$$

provided $f(x) > 0$ for $x$ near $a$, if $n$ is even.
\endthm

\begthm{2.4}{Limits of Polynomial and Rational Functions}
Assume $p$ and $q$ are polynomials and $a$ is a constant.

\par a. Polynomial functions
$$\
\lim_{x \to a} p(x) = p(a)
$$


\par b. Rational functions

$$\
\lim_{x \to a} {p(x) \over q(x)} = {p(a) \over q(a)}
$$

provided $q(a) \ne 0$.
\endthm

\begthm{2.3h}{Limit laws for One-Sided Limits}
Laws 2.3a-2.3f hold with $\lim\limits_{x \to a}$ replaced with
$\lim\limits_{x \to a^+}$ or $\lim\limits_{x \to a^-}$. Law 2.3g is
modified as follows.

\par a.

$$\
\lim_{x \to a^+}(f(x))^{1/n} =
\bigg(
\lim_{x \to a^+} f(x)
\bigg)^{1/n}
$$

provided $f(x) \ge 0$, for $x$ near $a$ with $x > a$, if $n$ is even.

\par b.

$$\
\lim_{x \to a^-}(f(x))^{1/n} =
\bigg(
\lim_{x \to a^-} f(x)
\bigg)^{1/n}
$$

provided $f(x) \ge 0$, for $x$ near $a$ with $x < a$, if $n$ is even.

\endthm

\begthm{2.5}{The Squeeze Theorem}
Assume the functions $f$, $g$, and $h$ satisfy $f(x) \le g(x) \le h(x)$
for all values of $x$ near $a$, except possibly at $a$.
If $\lim\limits_{x \to a} f(x) = \lim\limits_{x \to a} h(x) = L$,
then $\lim\limits_{x \to a} g(x) = L$.
\endthm

\begthm{2.6}{Limits at Infinity of Powers and Polynomials}
Let $n$ be a positive integer and let $p$ be the polynomial

$p(x) = a_n x^n + a_{n - 1}x^{n - 1} + \cdots + a_2 x^2 +
a_1x + a_0$, where $a_n \ne 0$.
\par 1.
$\lim\limits_{x \to \pm\infty} x^n = \infty$ when $n$ is even.

\par 2.
$\lim\limits_{x \to \infty} x^n = \infty$
and
$\lim\limits_{x \to -\infty} x^n = -\infty$
when $n$ is odd.

\par 3.
$\lim\limits_{x \to \pm\infty} {1 \over x^n} =
\lim\limits_{x \to \pm\infty} {x^{-n}} = 0
$.

\par 4.
$\lim\limits_{x \to \pm\infty} p(x)=
\lim\limits_{x \to \pm\infty} a_nx^n= \pm\infty$,
depending on the degree of the polynomial and the sign
of the leading coefficient $a_n$.

\endthm

\begthm{2.7}{End Behavior and Asymptotes of Rational Functions}
Suppose $f(x) = {p(x) \over q(x)}$ is a rational function,
where
$$\
p(x) = a_mx^m + a_{m-1}x^{m-1} + \cdots + a_2x^2 + a_0
$$

and

$$\
q(x) = b_nx^n + b_{n-1}x^{n-1} + \cdots + b_2x^2 + b_0
$$

with $a_m \ne 0$ and $b_n \ne 0$.

\par {\bf a. Degree of numerator less than degree of denominator.}
If $m < n$, then $\lim\limits_{x \to \pm\infty} f(x) = 0$,
and $y = 0$ is a horizontal asymptote of $f$.

\par {\bf b. Degree of numerator equals degree of denominator.}
If $m = n$, then $\lim\limits_{x \to \pm\infty} f(x) = a_m/b_n$,
and $y = a_m/b_n$ is a horizontal asymptote of $f$.

\par {\bf c. Degree of numerator greater than degree of denominator.}
If $m > n$, then $\lim\limits_{x \to \pm\infty} f(x) = \infty$ or $-\infty$,
and $f$ has no horizontal asymptote.

\par {\bf d. Slant asymptote.}
If $m = n + 1$, then $\lim\limits_{x \to \pm\infty} f(x) = \infty$
or $-\infty$, and $f$ has no horizontal asymptote, but $f$
has a slant asymptote.

\par {\bf e. Vertifcal asymptotes.} Assuming $f$ is in
reduced form ($p$ and $q$ share no common factors), vertical
asymptotes occur at the zeros of $q$.
\endthm

\begthm{2.8}{End Behavior of $e^x$, $e^{-x}$, and $\ln x$}
The end behavior for $e^x$ and $e^{-x}$ on $(-\infty, infty)$
and on $\ln x$ on $(0, \infty)$ is given by the following
limits:

$$\
\lim_{x \to \infty} e^x = \infty,
\lim_{x \to -\infty} e^x = 0
$$

$$\
\lim_{x \to \infty} e^{-x} = 0,
\lim_{x \to -\infty} e^{-x} = \infty
$$

$$\
\lim_{x \to 0^+} \ln x = -\infty,
\lim_{x \to \infty} \ln x = \infty
$$
\endthm

\begthm{2.9}{Continuity Rules}
If $f$ and $g$ are continuous at $a$, then the following
functions are also continuous at $a$. Assume $c$ is a constant
and $n > 0$ is an integer.

\par {\bf a.} $f + g$

\par {\bf b.} $f - g$

\par {\bf c.} $cf$

\par {\bf d.} $fg$

\par {\bf e.} $f/g$, provided $g(a) \ne 0$

\par {\bf f.} $(f(x))^n$

\endthm

\begthm{2.10}{Polynomial and Rational Functions}
\par {\bf a.} A polynomial function is continuous for all x.

\par {\bf b.} A rational function (a function of the form
${p \over q}$, where $p$ and $q$ are polynomials)
is continuous for all $x$ for which $q(x) \ne 0$.

\endthm

\begthm{2.11}{Continuity of Composite Functions at a Point}
if $g$ is continuous at $a$ and $f$ is continuous at $g(a)$,
then the composite function $g \circ f$ is continuous at $a$.
\endthm

\begthm{2.12}{Limits of Composite Functions}
\par {\bf 1.} if $g$ is continuous at $a$ and $f$ is continuous
at $g(a)$, then

$$\
\lim_{x \to a} f(g(x)) = f\big(\lim_{x \to a} g(x)\big).
$$

\par {\bf 2.} if $g(x) = L$ and $f$ is continuous at $L$,
then

$$\
\lim_{x \to a} f(g(x)) = f\big(\lim_{x \to a} g(x)\big).
$$

\endthm

\begthm{2.13}{Continuity of Functions with Roots}
Assume $n$ is a positive integer. If $n$ is an odd integer,
then $(f(x))^{1/n}$ is continuous at all points at which
$f$ is continuous.
\par
If $n$ is even, then $(f(x))^{1/n}$ is continuous at all points
$a$ at which $f$ is continuous and $f(a) > 0$.
\endthm

\begthm{2.14}{Continuity of Inverse Functions}
If a function $f$ is continuous on an interval $I$ and
has an inverse on $I$, then its inverse $f^{-1}$ is also
continuous (on the interval consisting of the points
$f(x)$, where $x$ is in $I$).
\endthm

\begthm{2.15}{Continuity of Transcendental functions}
TODO.
\endthm

\begthm{2.16}{Intermediate Value Theorem}
Suppose $f$ is continuous on the interval $[a, b]$ and $L$
is a number strictly between $f(a)$ and $f(b)$. Then there
exists at lest one number $c$ in $(a, b)$ satisfying $f(c) = L$.
\endthm

\begthm{3.10}{Trigonometric Limits}
$$\
\lim_{x \to 0} = {\sin x \over x} = 1
$$

$$\
\lim_{x \to 0} = {\cos x - 1 \over x} = 0
$$
\endthm

\begthm{3.11}{Derivatives of Sine and Cosine}
$$\
{d \over dx}(\sin x) = \cos x
$$

$$\
{d \over dx}(\cos x) = -\sin x
$$
\endthm

\begthm{3.12}{Derivatives of the Trigonometric Functions}
$$\
{d \over dx}
(\sin x)
=
\cos x
$$

$$\
{d \over dx}
(\cos x)
=
-\sin x
$$

$$\
{d \over dx}
(\tan x)
=
\sec^2 x
$$

$$\
{d \over dx}
(\cot x)
=
-\csc^2 x
$$

$$\
{d \over dx}
(\sec x)
=
\sec x \tan x
$$

$$\
{d \over dx}
(\csc x)
=
-\csc x \cot x
$$
\endthm

\begthm{3.13}{The Chain Rule}
Suppose $y = f(u)$ is differentiable at $u = g(x)$ and $u = g(x)$
is differentiable at $x$. The composite function $f(g(x))$ is
differentiable at $x$, and its derivative can be expressed
in two equivalent ways.

$$\
{dy \over dx} = {dy \over du} \cdot {du \over dx}
$$

$$\
{d \over dx}(f(g(x))) = f'(g(x)) \cdot g'(x)
$$
\endthm

\begthm{3.14}{Chain Rule for Powers}
If $g$ is differentiable for all $x$ in its domain and $p$
is a real number, then

$$\
{d \over dx} \big((g(x))^p\big) = p(g(x))^{p - 1}g'(x)
$$
\endthm

\begthm{3.15}{Derivative of $\ln x$}
$$\
{d \over dx} (\ln x) = {1 \over x}, \hbox{for} \, x > 0
$$

$$\
{d \over dx} (\ln |x|) = {1 \over x}, \hbox{for} \, x \ne 0
$$

If $u$ is differentiable at $x$ and $u(x) \ne 0$, then

$$\
{d \over dx} (\ln |u(x)|) = {
u'(x) \over u(x)
}
$$
\endthm

\begthm{3.16}{Theorem of $b^x$}
If $b > 0$ and $b \ne 1$, then for all $x$,

$$\
{d \over dx}(b^x) = b^x \ln b
$$
\endthm

\begthm{3.17}{General Power Rule}
For real numbers $p$ and for $x > 0$,
$$\
{d \over dx}(x^p) = px^{p - 1}.
$$

Furthermore, if $u$ is a positive differentiable function
on its domain, then

$$\
{d \over dx} (u(x)^p) = p(u(x))^{p - 1} \cdot u'(x)
$$
\endthm

\begthm{3.18}{Derivative of $\log_{b}x$}
If $b > 0$ and $b \ne 1$, then
$$\
{d \over dx} (log_{b} x) = {1 \over {x \ln b}}, \hbox{for}
\, x > 0
$$

and

$$\
{d \over dx}(log_{b}|x|) = {1 \over {x \ln b}}, \hbox{for}
\, x \ne 0
$$

\endthm

\begthm{3.19}{Derivative of Inverse Sine}
$$\
{d \over dx}(\sin^{-1} x) = {
1
\over
\sqrt{1 - x^2}
}, \, \hbox{for} -1 < x < 1
$$
\endthm

\begthm{3.20}{Derivatives of Inverse Trigonometric Functions}
$$\
{d \over dx}
(\sin^{-1} x)
=
{
1
\over
\sqrt{1 - x^2}
}
\qquad
{d \over dx}
(\cos^{-1} x)
=
-{
1
\over
\sqrt{1 - x^2}
},
\, \hbox{for} \, -1 < x < 1
$$

$$\
{d \over dx}
(\tan^{-1} x)
=
{
1
\over
1 + x^2
}
\qquad
{d \over dx}
(\cot^{-1} x)
=
-{
1
\over
1 + x^2
},
\, \hbox{for} \, -\infty < x < \infty
$$

$$\
{d \over dx}
(\sec^{-1} x)
=
{
1
\over
|x|\sqrt{x^2 - 1}
}
\qquad
{d \over dx}
(\csc^{-1} x)
=
-{
1
\over
|x|\sqrt{x^2 - 1}
},
\, \hbox{for} \, |x| > 1
$$

\endthm

\begthm{3.21}{Derivative of the Inverse Function}
Let $f$ be differentiable and have an invese on an interval $I$.
If $x_0$ is a point of $I$ at which $f'(x_0) \ne 0$, then $f^{-1}$
is differentiable at $y_0 = f(x_0)$ and

$$\
(f^{-1})'(y_0) = {1 \over f'(x_0)},
\, \hbox{where} \, y_0 = f(x_0).
$$
\endthm

\begthm{4.1}{Extreme Value Theorem}
A function that is continuous on a cl,soed interval $[a, b]$
has an absolute maximum value and an absolute minimum value on
that interval.
\endthm

\begthm{4.2}{Local Extreme Value Theorem}
If $f$ has a local maximum or minimum value at $c$ and
$f'(c)$ exists, then $f'(c) = 0$.
\endthm

\begthm{4.3}{Rolle's Theorem}
Let $f$ be continuous on a clsoed interval $[a,b]$ and differentiable on $(a, b)$
with $f(a) = f(b)$. There is at least one point $c$ in $(a, b)$ such that
$f'(c) = 0$.
\endthm

\begthm{4.4}{Mean Value Theorem}
If $f$ is continuous on the closed interval $[a,b]$ and differentiable
on $(a, b)$, then there is at least one point $c$ in $(a, b)$
such that
$$\
{f(b) - f(a) \over b - a} = f'(c)
$$
\endthm

\begthm{4.5}{Zero Derivative Implies Constant Function}
If $f$ is differentiable and $f'(x) = 0$ at all points of an
open interval $I$, then $f$ is a constant function on $I$.
\endthm

\begthm{4.6}{Functions with Equal Derivatives Differ by a Constant}
If two functinos of the property that $f'(x) = g'(x)$, for
all $x$ of an open interval $I$, then $f(x) - g(x) = C$ on $I$,
where $C$ is a constant; that is, $f$ and $g$ differ by a constant.
\endthm

\begthm{4.7}{Test for Intervals of Increase and Decrease}
Suppose $f$ is continuous on an interval $I$ and differentiable
at all interior points of $I$. If $f'(x) > 0$ at all interior
points of $I$, then $f$ is increasing on $I$. If $f'(x) < 0$
at all interior points of $I$, then $f$ is decreasing on $I$.
\endthm

\begthm{4.8}{First Derivative Test}
Assume $f$ is continuous on an interval that contains a critical
point $c$, and assume $f$ is differentiable on an interval containing
$c$, except perhaps at $c$ itself.
\par
If $f'$ changes sign from positive to negative as $x$ increases
through $c$, then $f$ has a {\bf local maximum} at $c$.
\par
If $f'$ changes sign from netagive to positive as $x$ increases
through $c$, then $f$ has a {\bf local minimum} at $c$.
\par
If $f'$ is positive on both sides near $c$ or negative on boths
sides near $c$, then $f$ has no local extreme value at $c$.
\endthm

\begthm{4.9}{One Local Extremum Implies Absolute Extremum}
Suppose $f$ is continuous on an interval $I$ that contains
exactly one local extremum at $c$.
\par
If a local maximum occurs at $c$, then $f(c)$ is the absolute
maximum of $f$ on $I$.
\par
If a local minimum occurs at $c$, then $f(c)$ is the absolute
minimum of $f$ on $I$.
\endthm

\begthm{4.10}{Test for Concavity}
Suppose $f''$ exists on an open interval $I$.
\par
If $f'' > 0$ on $I$, then $f$ is concave up on $I$.
\par
If $f'' < 0$ on $I$, then $f$ is concave down on $I$.
\par
If c is a point of $I$ at which $f''$ changes sign at $c$
(from positive to negative, or vice versa), then $f$ has an
inflection point at $c$.
\endthm

\begthm{4.11}{Second Derivative Test for Local Extrema}
Suppose $f''$ is continuous on an open interval containing $c$
with $f'(c) = 0$.
\par
If $f''(c) > 0$, then $f$ has a local minimum at $c$.
\par
If $f''(c) < 0$, then $f$ has a local maximum at $c$.
\par
If $f''(c) = 0$, then the test is nonconclusives; $f$
may have a local maximum, a local minimum, or neither at $c$.
\endthm

\bye
